---
output: github_document
bibliography: ./inst/REFERENCES.bib
link-citations: true
csl: ./inst/apa.csl
css: ./inst/styles.css
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  citation = "color: lightblue;",
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width="100%"
)
```

# PIFDataModel: Simplifying Structured Data Management for Pathogens in Food Database

## Overview

The **PIFDataModel** R package is an integral component of the Pathogens in Foods Database Project [(PIF: https://pif.esa.ipb.pt/)](https://pif.esa.ipb.pt/), a comprehensive initiative to facilitate the sharing, analysis, and usage of structured data and JSON schemas related to the **PIF** Database. This package is designed to support researchers, food safety professionals, and decision-makers in accessing and working with well-structured datasets and schemas for analyzing data on foodborne pathogens.

## Key Features

- **Access Predefined Schemas:** Includes predefined JSON schemas for organizing and validating data, such as bacteria count data.
- **Validation Tools:** Provides tools to validate JSON data against schemas (future development).
- **Data Management:** Simplifies the integration of structured data into R workflows.
- **Documentation and Resources:** Offers extensive documentation to facilitate the use of JSON schemas in research projects.

## Installation

Install the development version of **PIFDataModel** directly from GitHub:

```{r install, eval=FALSE}
# Install the remotes package if not already installed
install.packages("remotes")

# Install the PIFDataModel package from GitHub
remotes::install_github("fsqanalytics/PIFDataModel")
```

## Contributing

Contributions are welcome! If you'd like to contribute to the development of this package, please visit the [GitHub repository](https://github.com/fsqanalytics/PIFDataModel) to create issues or submit pull requests.

## Acknowledgments

The [PIF Database](https://pif.esa.ipb.pt/) and the [PIFDataModel](https://github.com/fsqanalytics/PIFDataModel) package are developed as part of a research project funded by the [European Food Safety Authority (EFSA)](https://www.efsa.europa.eu/en). We extend our gratitude to all contributors and collaborators who have supported this initiative.

For more information about the project, please read the paper:
<span style="color:blue">@Gonzales-BarronMRA2025</span>.

## References
