---
title: "Using PIFDataModel: Managing and Validating JSON Schemas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using PIFDataModel: Managing and Validating JSON Schemas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The `PIFDataModel` package provides tools to manage, validate, and document JSON schemas in R. This vignette demonstrates how to:
1. Load a JSON schema from the package.
2. Validate JSON data against the schema.
3. Generate documentation for a schema.

The package is designed to facilitate the integration and validation of structured JSON data in R workflows.

## Loading a Schema

The first step is to load a schema using the `load_schema()` function. This function reads a JSON schema file from the package's `inst/schemas/` directory and converts it into an R object.

```{r, eval=FALSE}
# Load a schema
schema <- load_schema("user_schema.json")
print(schema)
```

## Validating JSON Data

Use the validate_schema() function to check if a JSON object conforms to a specific schema. The input can be a JSON string or an R object (converted to JSON).

```{r, eval=FALSE}
# JSON data as a string
json_data <- '{"id": 1, "name": "Alice", "email": "alice@example.com"}'

# Validate the JSON data against the schema
validate_schema(json_data, "user_schema.json")
```

## Generating Documentation

The document_schema() function creates a human-readable description of a schema in markdown format. This is useful for sharing or documenting the structure of your schemas.

```{r, eval=FALSE}
# Generate markdown documentation for the schema
doc <- document_schema("user_schema.json")
cat(doc)
```


## User Schema

This schema validates user objects.

### Properties

- `id`: Integer, required.
- `name`: String, required.
- `email`: String, must be a valid email.






