---
title: "Structures in PIF: Standardized JSON for Biological Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structures in PIF: Standardized JSON for Biological Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The `PIFDataModel` package is designed to standardize data storage and handling for biological studies, particularly focusing on structured JSON schemas. This vignette provides an overview of the available data structures, their purposes, and examples of how to access and utilize them.

# Structures Categories

Below is an overview of the structures included in the `PIFDataModel` package:

| **Category**        | **Description**                                              | **Link**                                                        |
|---------------------|--------------------------------------------------------------|--------------------------------------------------------------|
| **Bacteria**        | Contains the count data for bacteria                          | [View](../../inst/extdata/structs/bacteria_count_results.json)  |
|                     | Contains the essays related to bacteria research              | [View](../../inst/extdata/structs/bacteria_essays.json)        |
|                     | Stores general results from bacteria-related studies          | [View](../../inst/extdata/structs/bacteria_general_results.json)|
|                     | Contains prevalence results for bacteria                      | [View](../../inst/extdata/structs/bacteria_prevalence_results.json)|
| **FAO**             | Contains information about FAO areas                         | [View](../../inst/extdata/structs/fao_areas.json)              |
| **Food**            | Contains characteristics of food                              | [View](../../inst/extdata/structs/food_characteristics.json)   |
|                     | Describes the structure of different food types               | [View](../../inst/extdata/structs/food_structure.json)         |
| **Nematodes**       | Contains count results for nematodes                          | [View](../../inst/extdata/structs/nematodes_count_results.json)|
|                     | Contains essays related to nematodes                          | [View](../../inst/extdata/structs/nematodes_essays.json)       |
|                     | Provides food-related information for nematodes               | [View](../../inst/extdata/structs/nematodes_food_info.json)    |
|                     | Contains general results from nematode studies                | [View](../../inst/extdata/structs/nematodes_general_results.json)|
|                     | Contains essay data for nematode identification               | [View](../../inst/extdata/structs/nematodes_id_essay.json)     |
|                     | Contains essay data for nematode infectivity                  | [View](../../inst/extdata/structs/nematodes_infec_essay.json)  |
|                     | Contains results related to nematode infection                | [View](../../inst/extdata/structs/nematodes_infec_results.json)|
|                     | Contains prevalence results for nematodes                     | [View](../../inst/extdata/structs/nematodes_prev_results.json) |
|                     | Contains essay data on nematode sample preparation            | [View](../../inst/extdata/structs/nematodes_sample_prep_essay.json)|
|                     | Defines the subtype tree for nematodes                        | [View](../../inst/extdata/structs/nematodes_subtype_tree.json) |
|                     | Describes the nematode types in a hierarchical structure      | [View](../../inst/extdata/structs/nematodes_type.json)         |
| **Protozoans**      | Contains enumeration results for protozoans                   | [View](../../inst/extdata/structs/protozoans_enum_results.json)|
|                     | Contains essays related to protozoans                         | [View](../../inst/extdata/structs/protozoans_essays.json)      |
|                     | Stores general results from protozoan studies                 | [View](../../inst/extdata/structs/protozoans_general_results.json)|
|                     | Contains essay data for protozoan identification              | [View](../../inst/extdata/structs/protozoans_id_essay.json)    |
|                     | Contains essay data for protozoan infectivity                 | [View](../../inst/extdata/structs/protozoans_infec_essay.json) |
|                     | Contains results related to protozoan infection               | [View](../../inst/extdata/structs/protozoans_infec_results.json)|
|                     | Contains prevalence results for protozoans                    | [View](../../inst/extdata/structs/protozoans_prev_results.json)|
|                     | Contains essay data on protozoan sample preparation           | [View](../../inst/extdata/structs/protozoans_sample_prep_essay.json)|
|                     | Defines the subtype classification for protozoans             | [View](../../inst/extdata/structs/protozoans_subtype.json)     |
|                     | Describes the protozoan types in a hierarchical structure     | [View](../../inst/extdata/structs/protozoans_type.json)        |
|                     | Lists the measurement units used for protozoan data           | [View](../../inst/extdata/structs/protozoans_units.json)       |
| **Study**           | Contains structural data for study schema                    | [View](../../inst/extdata/structs/study_structure.json)        |
| **Virus**           | Contains enumeration results for viruses                      | [View](../../inst/extdata/structs/virus_enum_results.json)     |
|                     | Contains essays related to virus research                     | [View](../../inst/extdata/structs/virus_essays.json)           |
|                     | Stores general results from virus studies                     | [View](../../inst/extdata/structs/virus_general_results.json)  |
|                     | Contains essay data for virus identification                  | [View](../../inst/extdata/structs/virus_id_essay.json)         |
|                     | Contains essay data for virus infectivity                     | [View](../../inst/extdata/structs/virus_infec_essay.json)      |
|                     | Contains results related to virus infection                   | [View](../../inst/extdata/structs/virus_infec_results.json)    |
|                     | Contains results for virus quality and molecular analysis     | [View](../../inst/extdata/structs/virus_quality_mol_results.json)|
|                     | Defines the subtype classification for viruses                | [View](../../inst/extdata/structs/virus_subtype.json)          |
|                     | Describes the virus types in a hierarchical structure         | [View](../../inst/extdata/structs/virus_type.json)             |

# Examples

Here are examples of how to load and work with the structures in R:

## Loading the `PIFDataModel` Package

```{r lib}
# Load the PIFDataModel package
library(PIFDataModel)
```

## Accessing the Bacteria Count Data Struct

```{r bact}
library(jsonlite)
# Load the JSON struct
bacteria_struct <- jsonlite::read_json("../inst/extdata/structs/bacteria_count_results.json")

# Print the struct
print(bacteria_struct)
```

### Example Output

Here is a sample of the `bacteria_count_data.json` struct:

<details>
<summary>Bacteria Count Data Struct</summary>
<pre>
{
  "species": "Escherichia coli",
  "count": 12345,
  "location": "River A",
  "date": "2025-01-22"
}
</pre>
</details>

# Additional Documentation

For more details on how to use the extdata/structs or integrate them into your workflows, visit the [`PIFDataModel` documentation](https://example.com/documentation).